---
title: "Regression Models Project"
author: "E.T. Moseley"
date: "January 28, 2017"
output: pdf_document
---

#Executive Summary

In this analysis, we will generate two models to explain differences in mileage per gallon using the mtcars data set. Our first linear model includes only the transmission to predict mileage per gallon. Our second linear fit includes weight in addition to the transmission type.

Furthermore, we have determined that a manual transmission is associated with an increase in gas mileage on average of 7.25 mpg (disregarding all other variables).

#Exploratory Data Analysis

In this analysis, we will be exploring the "mtcars" dataset to determine what variables are associated with increased or decreased mileage per gallon. Precisely, we will be investigating what relationship (if any) the transmission, automatic or manual, has on the miles per gallon.

Per ?mtcars documentation:

A data frame with 32 observations on 11 variables.

[, 1] mpg	Miles/(US) gallon

[, 2]	cyl	Number of cylinders

[, 3]	disp	Displacement (cu.in.)

[, 4]	hp	Gross horsepower

[, 5]	drat	Rear axle ratio

[, 6]	wt	Weight (lb/1000)

[, 7]	qsec	1/4 mile time

[, 8]	vs	V/S

[, 9]	am	Transmission (0 = automatic, 1 = manual)

[,10]	gear	Number of forward gears

[,11]	carb	Number of carburetors


```{r}
attach(mtcars)#Load data

str(mtcars)
```

```{r, echo=FALSE}
#pairs(mtcars, main = "mtcars")
```

Let's see if there may be a relationship between miles per gallon and transmission:

```{r}
auto <- mtcars[(mtcars$am == 0),]#0 = automatic
manu <- mtcars[(mtcars$am == 1),]#1 = manual
summary(auto$mpg)
summary(manu$mpg)
```

Manual transmissions have higher Median and Mean mileage per gallon, with Medians of 22.80 vs 17.30 (manual vs automatic) and Means of 24.39 vs 17.15.

#Fitting Multiple Models

We can run a linear regression with mpg predicted by transmission type to determine if the relationship has statistical significance.

```{r}
fit <- lm(mpg ~ am, data = mtcars)
summary(fit)$coefficients
```

The probability that the transmission is not significant to mileage per gallon is 2.85e-04, which is much less than zero. Furthermore, we see that the slope of this model, 7.245 (the difference in mileage associated with a manual vs an automatic transmission) is precisely identical to the difference in means of both respective data:

```{r}
mean(manu$mpg) - mean(auto$mpg)
```

It is safe to reject the null hypothesis and investigate further. We will also begin to explore if weight affects gas mileage:

```{r}
coplot(mpg ~ wt | as.factor(am), data = mtcars, panel = panel.smooth, rows = 1)
```

Visually, it would appear that, for both manual and automatic transmissions, weight affects fuel mileage.

We will investigate briefly if weight is a factor by adding it to our regression model and seeing if it improves the fit.

```{r}
fit2 <- lm(mpg ~ am + wt, data = mtcars)
summary(fit2)$coefficients
```

Here, weight seems to be more predictive of mileage per gallon than does the transmittion type, with P-values of 1.87e-07 and 9.88e-01, respectively.

We can now test to determine if these two models are distinct.

```{r}
anova(fit, fit2, test="Chisq")
```

In this case, with the addition of wt to the model, the probability that the variation in predictions in either model is due only to chance is 1.12e-11, which is insignificant.

#Apprendix
##Residual plots for fit (mpg ~ am) and fit2 (mpg ~ am + wt)

```{r}
par(mfrow=c(2,2))
plot(fit)

plot(fit2)
```